{\bf Practical Concurrency Testing} \\ \normalsize \vspace{1em} {\em or, How I Learned to Stop Worrying and Love the Exponential Explosion}

### 2017-03-01 proposal/overview.tex
add HTM example in overview

Transactional memory (TM) may be implemented either in hardware, using special instructions and existing cache coherence algorithms,
or in software, via library calls and a log-based commit approach.
Software transactions (STM) \cite{stm-pldi06} can be used on any commodity processor, but must impose runtime overhead associated with logging.
Hardware transactions (HTM) \cite{htm-experience, htm-performance} achieve better performance by reusing existing cache coherence logic to detect conflicts, but require explicit support from the CPU, which is not yet widespread.
Haswell \cite{htm-haswell} is the first x86 architecture to support HTM,
offering three new instructions: \texttt{xbegin}, \texttt{xend}, and \texttt{xabort}, to begin, commit, and fail a transaction, respectively.

>>>>

Transactional memory (TM) may be implemented either in hardware, using special instructions and existing cache coherence algorithms,
or in software, via library calls and a log-based commit approach.
Software transactions (STM) \cite{stm-pldi06} can be used on any commodity processor, but must impose runtime overhead associated with logging.
Hardware transactions (HTM) \cite{htm-experience, htm-performance} achieve better performance by reusing existing cache coherence logic to detect conflicts, but require explicit support from the CPU, which is not yet widespread.
Haswell \cite{htm-haswell} is the first x86 architecture to support HTM,
offering three new instructions: \texttt{xbegin}, \texttt{xend}, and \texttt{xabort}, to begin, commit, and fail a transaction, respectively.
The example program in Figure~\ref{fig:htm-example} demonstrates how these instructions can be used to synchronize a simple shared access without locking overhead in the common case,
using GCC's compiler intrinsics \cite{htm-gcc}.

\begin{figure}[h]
	\begin{center}
		\begin{tabular}{l}
		\textbf{(All threads)} \\
		\hline
		\texttt{if ((status = \_xbegin()) == \_XBEGIN\_STARTED) \{} \\
		\texttt{~~~~if (too\_much(foo)) \{} \\
		\texttt{~~~~~~~~\_xabort();} \\
		\texttt{~~~~\} else \{} \\
		\texttt{~~~~~~~~foo++;} \\
		\texttt{~~~~~~~~\_xend();} \\
		\texttt{~~~~\}} \\
		\texttt{\} else if (status != \_XABORT\_EXPLICIT) \{} \\
		\texttt{~~~~mutex\_lock(\&m);} \\
		\texttt{~~~~foo++;} \\
		\texttt{~~~~mutex\_unlock(\&m);} \\
		\texttt{\}} \\
		\end{tabular}
	\end{center}
	\caption{An example program demonstrating the various HTM primitives.
		If the transaction in the top branch aborts,
		whether from a conflict or from the programmer's intention,
		execution will revert to the return of {\tt \_xbegin},
		and {\tt status} will be assigned an error code indicating the abort reason.
		The programmer can then use explicit synchronization, such as a mutex, to resolve the conflict.}
	\label{fig:htm-example}
\end{figure}
